% Related Work

Our work draws on three distinct areas of research: tabular synthetic data generation, blockchain-based provenance systems, and methods for measuring privacy and fairness in machine learning.

\textbf{Tabular data synthesis with GANs.}
Following Goodfellow et al.'s original adversarial training framework~\cite{goodfellow2014generative}, many GAN architectures have been modified for structured tabular data. CTGAN~\cite{xu2019ctgan} uses mode-specific normalisation and conditional training-by-sampling to model columns containing both continuous and categorical values. It is widely used as a baseline for tabular synthesis. CTAB-GAN+~\cite{zhao2022ctabgan} adds an auxiliary classifier to improve column-level accuracy. For formal privacy guarantees, PATE-GAN~\cite{jordon2018pate} integrates the Private Aggregation of Teacher Ensembles into its discriminator, and DP-CGAN~\cite{torkzadehmahani2020dp} applies carefully calibrated noise during gradient descent to satisfy a specific $(\varepsilon,\delta)$-differential-privacy budget. Sarmin et al.~\cite{sarmin2024privacy} evaluated these approaches and noted that single quality metrics cannot adequately measure both privacy and utility in synthetic data. Their observation directly informed the multi-metric design used in our system.

\textbf{Blockchain-backed data provenance.}
Hyperledger Fabric~\cite{androulaki2018hyperledger} is a widely adopted permissioned ledger that supports pluggable consensus mechanisms, including Raft~\cite{ongaro2014search} and PBFT~\cite{castro1999practical}, as well as channel-level data isolation. Kuo et al.~\cite{kuo2017blockchain} demonstrated how to use ledgers to track patient record provenance without exposing the records themselves. Veeraragavan et al.~\cite{veeraragavan2024blockchain} proposed a permissioned marketplace allowing consumers to rate synthetic data generators after testing the data. Their trust evaluation is subjective and occurs post-distribution. Our approach differs structurally: (i)~we conduct objective, metric-driven evaluations \emph{before} writing the commit to the ledger, and (ii)~we use a median consensus algorithm to aggregate independent examiner results into a verifiable quality certificate.

\textbf{Privacy and fairness measurement.}
Sweeney's $k$-anonymity framework~\cite{sweeney2002k} and the membership inference attack presented by Shokri et al.~\cite{shokri2017membership} measure re-identification risk using different methods---the former groups quasi-identifiers, while the latter uses adversarial model training. Yao et al.~\cite{yao2025dcr} showed that the common distance-to-closest-record (DCR) metric can be unreliable when datasets are tightly clustered, which further supports the need for using multiple evaluation metrics. For fairness, Hardt et al.~\cite{hardt2016equality} established equalised odds to evaluate group parity, Feldman et al.~\cite{feldman2015certifying} applied the four-fifths rule for compliance audits, and Calmon et al.~\cite{calmon2017optimized} designed pre-processing techniques to reduce bias before training. Our system builds on these methods by applying group rebalancing and selection-rate alignment after the data generation step.

\textbf{Gap in the existing literature.}
Table~\ref{tab:comparison} outlines the differences in functionality. To our knowledge, no existing system provides all three of these capabilities: an objective pre-commit verification process using multiple metrics, an immutable consensus-based record on a permissioned ledger, and an integrated evaluation covering fairness metrics and regulatory compliance.

\begin{table}[h]
\centering
\caption{Feature comparison of existing approaches and our framework.}
\begin{tabular}{lccc}
\toprule
\textbf{Capability} & \textbf{Central.} & \textbf{Reput.} & \textbf{Ours} \\
\midrule
Pre-commit verification & \checkmark & $\times$ & \checkmark \\
Objective quality metrics & \checkmark & $\times$ & \checkmark \\
Tamper-proof recording & $\times$ & \checkmark & \checkmark \\
Multi-party consensus & $\times$ & $\times$ & \checkmark \\
Fairness audit & $\times$ & $\times$ & \checkmark \\
Regulatory compliance map & Partial & $\times$ & \checkmark \\
\bottomrule
\end{tabular}
\label{tab:comparison}
\end{table}
